Studying public transit API query logs\\ to get an indication of travel flows

### 2016-01-03 paper.tex
dot
align center doesn't work
Temporally resolved error returned by "#" & "Package pgfkeys Error: The key '/tikz/text' requires a value."

\begin{figure}
\centering
\includegraphics[width=8.6cm,align=center]{querylogs}
\caption{Amount of queries on the route planning part of the iRail API between 2012 and 2016. In April 2015 an official app of the Belgian railways was discontinued, which explains the sudden raise of iRail API queries.}
\label{fig:querylogs}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=8.6cm,align=center]{avg-all}
\caption{Distribution of iRail route planning API queries on average per day of the week between 2012 and 2016}
\label{fig:average}
\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=8.3cm,align=center]{dec21}
%\caption{Comparison of route planning queries on December 21st 2015 with the average on a Monday between 2012 and 2016. Whilst being a school holiday, there is no notable difference, except for the peek hours being a bit later, }
%\label{fig:dec21}
%\end{figure}

\begin{figure}
\centering
\includegraphics[width=8.6cm,align=center]{dec24}
\caption{Comparison of route planning queries on December 24th (Christmass eve) 2015 with the average on a Thursday between 2012 and 2016. The illustration shows that people started going home earlier than usual.}
\label{fig:dec24}
\end{figure}

@@

Visualisation is frequently used to make data analysis tangible, so that the results can be communicated and debated {Robinson, 2008 #121}. This is crucial for our research since the questions presented are exploratory in nature {Kraak, 2008 #150} and can be addressed in many ways. Movement data has received substantial attention from the visualisation community and a range of techniques has been developed to support analysis {Andrienko, 2012 #196}. Of these techniques, flow visualisations facilitate the comparison of aggregated movement over space and time. There are three types of flow visualisations: Line based representations preserves the complete trajectory while matrix type representations take only the start and end locations into account.

Line based representations indicate movement on a path with the shape of the line. The thickness of each line is scaled to the volume of flow along a path. The direction of flow is commonly indicated with an arrowhead. This representation is detailed and straightforward to understand but becomes cluttered when lines intersect. A number of solutions have been proposed for clutter reduction. Filtering is the most commonly practiced. The same goal can be achieved by clustering lines that have similar properties like common start, end or intermediate locations {Andrienko, 2007 #198}. Hierarchical {Guo, 2009 #199} and density based {Rinzivillo, 2008 #201} clustering have also been introduced for this purpose. Both solutions are effective at reducing clutter yet it should be noted that excessive aggregation or filtering removes a significant amount of information from the visualisation. Another solution to reduce clutter is to transform the visual representation. This is achieved by rerouting {Phan, 2005 #88} or bundling {Hurter, 2014 #200} lines in close proximity so that they appear grouped. Unlike clustering, this solution changes the shape of lines by way of interpolation to obtain an ideal layout. In most cases, the interpolation is regulated by conditions to avoid intersections and follow key geographic features such as roads, rivers or coastlines so that the layout appears natural. This solution emphasises major flows within the visualisation since large groups gain visual dominance. Nonetheless, there are two drawbacks to this solution. First, short distance flows that may be important at local scales are de-emphasised or lost in the transformation. Next, the lines do not depict actual paths of movement thus this solution may not be suitable for applications that require strict geographic accuracy.

Matrix type representations are frequently used when the origin and destination preside over intermediate locations on the path of movement. This type of representation is referred to as origin-destination (OD) matrix. Rows and columns correspond to locations while cells are coloured to express the volume of flow. OD matrices can be reordered to emphasise connectivity between locations but the lack of geographic context is a distinct disadvantage. Several solutions have been developed to address this shortcoming. These are based on the notion of small multiples: The matrix is arranged in a geographical order so that the cells correspond to locations on the map. Then, a smaller map of the geographical context is nested in each cell to enable more intuitive comparisons {Guo, 2006 #202}. The limitation of this solution is similar to other forms or small multiples in general - the layout space allocated to each nested map becomes smaller as the number of cells increase. In this instance, nesting detailed geographical maps is problematic since majority of the geographical features become illegible due to scaling. OD Map {Wood, 2010 #86} proposes to nests a series of geographically ordered matrices that are more expressive of spatial relations and remain legible even when a small amount of layout space allocated to each cell.

@@

\begin{figure}
\centering
\includegraphics[width=8.1cm,align=center]{brussels}
\caption{Pair of chord visualization of the city of Brussels: transit flows towards Brussels from elsewhere in Belgium and vice versa.}
\label{fig:brussels}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=8.1cm,align=center]{antwerp}
\caption{Pair of chord visualization of the city of Antwerp: transit flows towards Antwerp from elsewhere in Belgium and vice versa.}
\label{fig:antwerp}
\end{figure}

@@

In \cref{fig:LDFAxis2} we illustrate these two options as two extremes, with other options that are yet to be discovered.
When a \emph{rather low expressive} server only allows to set e.g., a departure station and a departure time, then the server cannot log the arrival station, yet the client is still able to plan a route by executing the algorithm on the client-side~\cite{lc}.
Again, we are not able to rely on the query logs

>>>>

\begin{figure}
\centering
\includegraphics[width=8.6cm]{querylogs}
\caption{Amount of queries on the route planning part of the iRail API between 2012 and 2016. In April 2015 an official app of the Belgian railways was discontinued, which explains the sudden raise of iRail API queries.}
\label{fig:querylogs}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=8.6cm]{avg-all}
\caption{Distribution of iRail route planning API queries on average per day of the week between 2012 and 2016}
\label{fig:average}
\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=8.3cm]{dec21}
%\caption{Comparison of route planning queries on December 21st 2015 with the average on a Monday between 2012 and 2016. Whilst being a school holiday, there is no notable difference, except for the peek hours being a bit later, }
%\label{fig:dec21}
%\end{figure}

\begin{figure}
\centering
\includegraphics[width=8.6cm]{dec24}
\caption{Comparison of route planning queries on December 24th (Christmass eve) 2015 with the average on a Thursday between 2012 and 2016. The illustration shows that people started going home earlier than usual.}
\label{fig:dec24}
\end{figure}

@@

Visualisation is frequently used to make data analysis tangible, so that the results can be communicated and debated \todo{Robinson, 2008 -121}. This is crucial for our research since the questions presented are exploratory in nature \todo{Kraak, 2008 -150} and can be addressed in many ways. Movement data has received substantial attention from the visualisation community and a range of techniques has been developed to support analysis \todo{Andrienko, 2012 -196}. Of these techniques, flow visualisations facilitate the comparison of aggregated movement over space and time. There are three types of flow visualisations: Line based representations preserves the complete trajectory while matrix type representations take only the start and end locations into account.

Line based representations indicate movement on a path with the shape of the line. The thickness of each line is scaled to the volume of flow along a path. The direction of flow is commonly indicated with an arrowhead. This representation is detailed and straightforward to understand but becomes cluttered when lines intersect. A number of solutions have been proposed for clutter reduction. Filtering is the most commonly practiced. The same goal can be achieved by clustering lines that have similar properties like common start, end or intermediate locations \todo{Andrienko, 2007 -198}. Hierarchical \todo{Guo, 2009 -199} and density based \todo{Rinzivillo, 2008 -201} clustering have also been introduced for this purpose. Both solutions are effective at reducing clutter yet it should be noted that excessive aggregation or filtering removes a significant amount of information from the visualisation. Another solution to reduce clutter is to transform the visual representation. This is achieved by rerouting \todo{Phan, 2005 -88} or bundling \todo{Hurter, 2014 -200} lines in close proximity so that they appear grouped. Unlike clustering, this solution changes the shape of lines by way of interpolation to obtain an ideal layout. In most cases, the interpolation is regulated by conditions to avoid intersections and follow key geographic features such as roads, rivers or coastlines so that the layout appears natural. This solution emphasises major flows within the visualisation since large groups gain visual dominance. Nonetheless, there are two drawbacks to this solution. First, short distance flows that may be important at local scales are de-emphasised or lost in the transformation. Next, the lines do not depict actual paths of movement thus this solution may not be suitable for applications that require strict geographic accuracy.

Matrix type representations are frequently used when the origin and destination preside over intermediate locations on the path of movement. This type of representation is referred to as origin-destination (OD) matrix. Rows and columns correspond to locations while cells are coloured to express the volume of flow. OD matrices can be reordered to emphasise connectivity between locations but the lack of geographic context is a distinct disadvantage. Several solutions have been developed to address this shortcoming. These are based on the notion of small multiples: The matrix is arranged in a geographical order so that the cells correspond to locations on the map. Then, a smaller map of the geographical context is nested in each cell to enable more intuitive comparisons \todo{Guo, 2006 -202}. The limitation of this solution is similar to other forms or small multiples in general - the layout space allocated to each nested map becomes smaller as the number of cells increase. In this instance, nesting detailed geographical maps is problematic since majority of the geographical features become illegible due to scaling. OD Map \todo{Wood, 2010 -86} proposes to nests a series of geographically ordered matrices that are more expressive of spatial relations and remain legible even when a small amount of layout space allocated to each cell.

@@

\begin{figure}
\centering
\includegraphics[width=8.1cm]{brussels}
\caption{Pair of chord visualization of the city of Brussels: transit flows towards Brussels from elsewhere in Belgium and vice versa.}
\label{fig:brussels}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=8.1cm]{antwerp}
\caption{Pair of chord visualization of the city of Antwerp: transit flows towards Antwerp from elsewhere in Belgium and vice versa.}
\label{fig:antwerp}
\end{figure}

@@

In \cref{fig:LDFAxis2} we illustrate these two options as two extremes, with other options that are yet to be discovered.
When a \emph{rather low expressive} server only allows to set e.g., a departure station and a departure time, then the server cannot log the arrival station, yet the client is still able to plan a route by executing the algorithm on the client-side~\cite{lc}.
Again, we are not able to rely on the query logs.
